---
title: 'Employee Data'
author: "Josephine Haag"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    highlight: default
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)

```

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
# Load ggplot2, dplyr, and all the other tidyverse packages
library(tidyverse)  
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(patchwork)
library(ggplot2)
library(plotly)
library(ggthemr)
ggthemr('dust')

```

# IBM HR Analytics: Employee satisfaction and income 

During my internship at an HR Tech start-up Leapsome (leapsome.com) I learned a lot about the importance of employee retention and employee satisfaction.  Employee engagement and meaningful feedback, for example, are two drivers of productive employees. This data set on past and current employees at a company can give insights about what can contribute to workforce attrition. After cleaning the data set with data about around 1500 employees, I will analyze several HR data and different relationships within. 

*Note: Scroll over the interactive graphs to see the exact amount of servings*


```{r loading_HR_dataset, include=FALSE}
hr_dataset <- read_csv( "datasets_1067_1925_WA_Fn-UseC_-HR-Employee-Attrition.csv")
glimpse(hr_dataset)

```

```{r cleaning_HR_dataset, include=FALSE}
# clean data set
hr_cleaned <- hr_dataset %>% 
  clean_names() %>% 
  mutate(
    education = case_when(
      education == 1 ~ "Below College",
      education == 2 ~ "College",
      education == 3 ~ "Bachelor",
      education == 4 ~ "Master",
      education == 5 ~ "Doctor"
    ),
    environment_satisfaction = case_when(
      environment_satisfaction == 1 ~ "Low",
      environment_satisfaction == 2 ~ "Medium",
      environment_satisfaction == 3 ~ "High",
      environment_satisfaction == 4 ~ "Very High"
    ),
    job_satisfaction = case_when(
      job_satisfaction == 1 ~ "Low",
      job_satisfaction == 2 ~ "Medium",
      job_satisfaction == 3 ~ "High",
      job_satisfaction == 4 ~ "Very High"
    ),
    performance_rating = case_when(
      performance_rating == 1 ~ "Low",
      performance_rating == 2 ~ "Good",
      performance_rating == 3 ~ "Excellent",
      performance_rating == 4 ~ "Outstanding"
    ),
    work_life_balance = case_when(
      work_life_balance == 1 ~ "Bad",
      work_life_balance == 2 ~ "Good",
      work_life_balance == 3 ~ "Better",
      work_life_balance == 4 ~ "Best"
    )
  ) %>% 
  select(age, attrition, daily_rate, department,
         distance_from_home, education,
         gender, job_role, environment_satisfaction,
         job_satisfaction, marital_status,
         monthly_income, num_companies_worked, percent_salary_hike,
         performance_rating, total_working_years,
         work_life_balance, years_at_company,
         years_since_last_promotion)

```

## Employee Retention

### How often do people leave the company (attrition)?
```{r company_attrition_rate}
attrition_rate <- filter(hr_cleaned %>% 
         group_by(attrition) %>% 
         summarise(n = count(attrition)) %>% 
         mutate(attrition_rate = (n*100)/sum(n)), attrition == "Yes") # create new column for the attrition rate %

attrition_rate 

```

The attrition rate is 16.1%.


Next, I will investigate the job satisfaction among the employees and see if this differs by gender.
```{r job_satisfaction}
# calculate the average job satisfaction 
job_satisfaction_gender <- hr_dataset %>% 
  group_by(Gender) %>% 
  summarize(mean_js = mean(JobSatisfaction)) 
         
job_satisfaction_gender

# plot average job satisfaction 
job_satisfaction_gender_plot <- job_satisfaction_gender %>%
  ggplot(aes(
    x = Gender, 
    y = mean_js, 
    fill = Gender)) +
  geom_col() +
  # facet_wrap(~Gender) +
  ylim(0, 5) +
  labs(
    title = "Job Satisfaction by Gender", 
    subtitle = "Men are slightly more satisfied with their jobs than women",
    y = "Average Job Satisfaction") + # add titles
  theme(legend.position = "none") # remove legend
  NULL

job_satisfaction_gender_plot 
```
Job satisfaction seems to be slightly higher for men than for women. 

I will continue by investigating how are job satisfaction and work-life-balance are distributed

```{r job_sat}
# find distribution of job satisfaction
job_sat <- hr_cleaned %>% 
  group_by(job_satisfaction) %>% 
  summarise(n = count(job_satisfaction)) %>% 
  mutate(job_sat_perc = round((n*100)/sum(n)))

job_sat

# reorder levels of job satisfaction 
job_sat$job_satisfaction <- factor(job_sat$job_satisfaction, c("Low", "Medium", "High", "Very High"))

# plot distribution
job_sat_plot <- job_sat %>%
  ggplot(aes(
    x = job_sat_perc, 
    y = job_satisfaction, 
    text = paste0("Count: ", n, ", percentage: ", job_sat_perc, "%"))) + # label for ggplotly
  geom_col(fill = "#fecc5c") +
  labs(
    title = "Distribution of Job Satisfaction", 
    y = "Level of Job Satisfaction",
    x = "Percentage of total") +
  NULL

# convert into interactive graph
job_sat_plotly <- ggplotly(job_sat_plot, tooltip = "text")

# add title for ggplotly
job_sat_plotly %>%
  layout(title = list(text = paste0('Distribution of Job Satisfaction',
                                    '<br>',
                                    '<sup>',
                                    'Most employees have a high or very high job satisfaction',
                                    '</sup>'))) # add title
```

```{r worklife_balance}
# find distribution of worklife balance
work_life_balance <- hr_cleaned %>% 
  group_by(work_life_balance) %>% 
  summarise(n = count(work_life_balance)) %>% 
  mutate(work_life_perc = round((n*100)/sum(n))) %>% 
  arrange(desc(work_life_perc))

work_life_balance

# reorder levels of work life balance 
work_life_balance$work_life_balance <- factor(work_life_balance$work_life_balance, c("Bad", "Good", "Better", "Best"))

# plot distribution
work_life_balance_plot <- work_life_balance %>%
  ggplot(aes(
    x = work_life_perc, 
    y = work_life_balance, 
    text = paste0("Count: ", n, ", percentage: ", work_life_perc, "%"))) + # label for ggplotly
  geom_col(fill = "#feb24c") +
  labs(
    title = "Distribution of Work Life Balance", 
    y = "Level of Work Life Balance",
    x = "Percentage of total") + # add titles
  NULL

# convert into interactive graph
work_life_balance_plotly <- ggplotly(work_life_balance_plot, tooltip = "text")

# add title for ggplotly
work_life_balance_plotly %>%
  layout(title = list(text = paste0('Distribution of Work Life Balance',
                                    '<br>',
                                    '<sup>',
                                    'Most employees have a "better" work life balance',
                                    '</sup>'))) 

```

Job satisfaction is right skewed, with around 60% of the values on the right.
Work life balance seems to better resemble a normal distribution, a bit skewed to the right.


## Monthly income

### Education and income

I will continue with investigating whether there is any relationship between monthly income and education as well as between monthly income and gender.
```{r monthly_income_education_plots, fig.width=20}
options(scipen=100, digits=2)

# reorder levels of education 
hr_cleaned$education <- factor(hr_cleaned$education, c("Below College", "College", "Bachelor", "Master", "Doctor"))

# calculate mean and median income per education level
monthly_income_ed <- hr_cleaned %>%
  group_by(education) %>%
  summarize(median_income = round(median(monthly_income)), mean_income = round(mean(monthly_income))) # round to nearest digit
  
monthly_income_ed

# plot median income per education level and include mean interactively
monthly_income_education_plot <- monthly_income_ed %>%
  ggplot(aes(
     x = education, 
     y = median_income,
     text = paste0("Median: $", median_income," mean: $", mean_income) # label for ggplotly (including mean income)
     )) + 
  geom_col(fill = "#fd8d3c") +
  labs(
    title = "Monthly income vs. education",
    x = "Education level",
    y = "Monthly income"
  ) +
  NULL

# convert into interactive graph
monthly_income_education_plotly <- ggplotly(monthly_income_education_plot, tooltip = "text")

# add title for ggplotly
monthly_income_education_plotly %>%
  layout(title = list(text = paste0('Income across eduation levels',
                                    '<br>',
                                    '<sup>',
                                    'Income increases with additional education level',
                                    '</sup>'))) 
```

Next we will plot the distribution of income by education level.

```{r distribution_by_education, fig.width=10}

monthly_income_education_distr <- hr_cleaned %>%
  ggplot(aes(
    x = monthly_income
  )) + 
  geom_histogram(fill = "#bd0026") + 
  facet_wrap(~education,ncol=2, scales = "free") + 
  labs(
    title = "Distribution of income by education",
    x = ""
  ) +
  NULL

monthly_income_education_distr

```

### Gender and income

```{r monthly_income_gender_plots, fig.width=20}
# calculate mean and median income per gender
monthly_income_sex <- hr_cleaned %>%
  group_by(gender) %>%
  summarize(median_income = round(median(monthly_income)), mean_income = round(mean(monthly_income))) # round to nearest digit
  
monthly_income_sex

# plot median income per gender and include mean interactively
monthly_income_gender_plot <- monthly_income_sex %>%
  ggplot(aes(
     x = gender, 
     y = median_income,
     text = paste0("Median: $", median_income," mean: $", mean_income) # label for ggplotly (including mean income)
     )) + 
  geom_col(fill = "#fc4e2a") +
  labs(
    title = "Monthly income per gender",
    x = "",
    y = "Monthly income"
  ) +
  NULL

# convert into interactive graph
monthly_income_gender_plotly <- ggplotly(monthly_income_gender_plot, tooltip = "text")

# add title for ggplotly
monthly_income_gender_plotly %>%
  layout(title = list(text = paste0('Income per gender',
                                    '<br>',
                                    '<sup>',
                                    'Median and mean income are higher for women',
                                    '</sup>'))) 

```
The monthly income increases with additional levels of education, which is intuitive. Interestingly, the median income is higher for employees with a College degree than for those with a Bachelor degree. The mean, however, is higher for the Bachelor degree. This might imply that a Bachelor is not worth much more than a College degree when it comes to monthly income.

When comparing income between gender, the female mean and median income is higher than the male ones. This can, however, be influenced by various factors, such as the job position and education levle.

### Job role and income

Next, we will investigate the relationship between income vs job role.
```{r income_jobrole_boxplot}
# plot income per job role
income_jobrole_boxplot <- hr_cleaned %>%
  ggplot(aes(
         x = monthly_income,
         y = reorder(job_role, monthly_income)
         )) + 
  geom_boxplot(fill="#e31a1c") +
  labs(
    title = "Income per job role",
    x = "Income",
    y = "Job role"
  ) +
  NULL

income_jobrole_boxplot

```
Managers have the highest income followed by research directors and then after a gap, healthcare representatives, manufacturing directors and sales executives. Sales representatives earn the least, however, they have a lot of outliers for this and other "low-paying" roles. 

### Age and income
1. Plot income vs age, faceted by `job_role`
```{r Income_vs_age_by_jobrole}

Income_vs_age_facet <- ggplot(hr_cleaned, aes(
    x = age, 
    y = monthly_income
  )) + 
  geom_col(width = 2, 
           fill = "#800026") + 
  facet_wrap(~job_role) +
  labs(
    title = "Income vs. age by job role",
    x = "Age",
    y = "Monthly income"
  ) +
  NULL

Income_vs_age_facet

```
### Job satisfaction vs income

```{r scatterplot}
# plot job satisfaction vs income
income_satisfaction_plot<- hr_cleaned %>%
  ggplot(aes(x = reorder(job_satisfaction, -monthly_income),
             y = monthly_income)) +
  geom_col(fill = "#6b2021") +
    labs(
    title = "Income vs. job satisfaction",
    subtitle = "Higher income relates to higher satisfaction",
    x = "Job satisfaction",
    y = "Monthly income"
  ) +
  NULL

income_satisfaction_plot

```
In conclusion, it seems that higher incomes relates to higer job satisfaction. The direction of causation is difficult to establish in this case. Employees might be more satisfied because they earn more. An alternative explanation is that those who are satisfied with their job also do a better job and, thus, earn more since they are being more productive.


