---
title: 'Alcohol consumption'
author: "Josephine Haag"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    highlight: zenburn
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---


```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(patchwork)
library(ggplot2)

```



# Where Do People Drink The Most Beer, Wine And Spirits?
```{r load_alcohol_data, include=FALSE}
library(fivethirtyeight)
data(drinks)

```



```{r glimpse_skim_data}
skim(drinks)
```

There are no missing values

```{r beer_plot}
# filter for top 25 beer consuming countries
top25_beer <- drinks %>% 
  top_n(25, beer_servings)  # rank top 25 by beer servings

top25_beer_plot <- top25_beer %>%
  ggplot(aes(
    x = beer_servings,
    y = reorder(country,beer_servings) # order countries by beer servings
  )) +
  geom_col(stat = "identity", width = 0.8, fill = "#feb24c") + # adapt bars and set a color to beer
  labs(
    x = "Beer Servings", 
    y = "", 
    title = "Top 25 Beer Consuming Countries", subtitle = "Namibia, Czech Republic, & Gabon drink most beer", caption = "Source: World Health Organization (GISAH), 2010") + # add labels to axis and title
  theme_minimal() +
  NULL

top25_beer_plot

```


```{r wine_plot}
# filter for top 25 wine consuming countries
top25_wine <- drinks %>% 
  top_n(25, wine_servings)  # rank top 25 by wine servings

top25_wine_plot <- top25_wine %>%
  ggplot(aes(
    x = wine_servings,
    y = reorder(country,wine_servings) # order countries by wine servings
  )) +
  geom_col(stat = "identity", width = 0.8, fill = "#a50f15") + # adapt bars and set a color to (red) wine
  labs(
    x = "Wine Servings", 
    y = "", 
    title = "Top 25 Wine Consuming Countries", subtitle = "France, Portugal, and Andorra drink most wine", caption = "Source: World Health Organization (GISAH), 2010") + # add labels to axis and title
  theme_minimal() +
  NULL

top25_wine_plot

```

```{r spirit_plot}
# filter for top 25 spirit consuming countries
top25_spirit <- drinks %>% 
  top_n(25, spirit_servings)  # rank top 25 by spirit servings

top25_spirit_plot <- top25_spirit %>%
  ggplot(aes(
    x = spirit_servings,
    y = reorder(country,spirit_servings) # order countries by spirit servings
  )) +
  geom_col(stat = "identity", width = 0.8, fill = "#896045") + # adapt bars and set a color to spirit (rum)
  labs(
    x = "Spirit Servings", 
    y = "", 
    title = "Top 25 Spirit Consuming Countries", subtitle = "Granada, Belarus, and Russia drink most spirit", caption = "Source: World Health Organization (GISAH), 2010") + # add labels to axis and title
  theme_minimal() +
  NULL

top25_spirit_plot

```

```{r patchwork}
# combine the three plots next to each other
top25_beer_plot + top25_wine_plot + top25_spirit_plot

```



> IMPLICATIONS

- In the wine consumption rankings, 9 of the top 10 are European countries. Most areas of Europe (especially the low latitudes) are very suitable for growing grapes due to sufficient sunlight and fertile soil. The typical ones are Southern France, Portugal, Andorra, Italy, and Greece. This may develop the people's habit of drinking wine. Compared to wine, the high beer consuming countries are mainly in higher latitudes (north of the Alps), since wheat is widely planted and is an important raw material for beer in these countries. In addition, the beer ranking includes the most continents among other liquors, which may implicate that beer is the most widely accepted and accessible alcohol in the world.
- In the spirits consumption ranking, six of the top 25 are island countries in the Caribbean and the Pacific, and 11 of the top 25 are East European countries. This conforms to the origins of two very famous spirit types - rum and vodka.


```{r}
# # install.packages(c("cowplot", "googleway", "ggplot2", "ggrepel", 
# # "ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata")
# # 
# # library("ggplot2")
# # theme_set(theme_bw())
# # library("sf")
# # library("rnaturalearth")
# # library("rnaturalearthdata")
# 
# 
# map.world <- map_data("world")
# 
# # INSPECT
# as.factor(drinks$country) %>% levels()
# 
# # RECODE NAMES
# # drinks$country <- recode(drinks$country
# #                                   ,'United States' = 'USA'
# #                                   ,'United Kingdom' = 'UK'
# #         )
# 
# head(map.world)
# 
# # LEFT JOIN
# map.world_joined <- left_join(map.world, drinks, by = c('region' = 'country'))
# 
# df.top_beer <- data.frame(country = c(top25_beer$country),stringsAsFactors = F)
# glimpse(df.top_beer)
# 
# #--------
# # GEOCODE
# #--------
# 
# geocode.country_points <- geocode(df.top_beer$country)
# 
# df.top_beer <- cbind(df.top_beer,geocode.top_beer)
# 
# # INSPECT
# print(df.top_beer)
# 
# 
# ggplot() +
#   geom_polygon(data = map.world_joined, aes(x = long, y = lat, group = group, fill = fill_flg)) +
#   geom_point(data = top_25_beer, aes(x = lon, y = lat), color = "#e60000") +
#   scale_fill_manual(values = c("#CCCCCC","#e60000")) +
#   labs(title = 'Countries with highest "talent competitiveness"'
#        ,subtitle = "source: INSEAD, https://www.insead.edu/news/2017-global-talent-competitiveness-index-davos") +
#   theme(text = element_text(family = "Gill Sans", color = "#FFFFFF")
#         ,panel.background = element_rect(fill = "#444444")
#         ,plot.background = element_rect(fill = "#444444")
#         ,panel.grid = element_blank()
#         ,plot.title = element_text(size = 30)
#         ,plot.subtitle = element_text(size = 10)
#         ,axis.text = element_blank()
#         ,axis.title = element_blank()
#         ,axis.ticks = element_blank()
#         ,legend.position = "none"
#         )
# 
# 
# 
# # ggplot(data = drinks) +
# #     geom_sf()
# # 
# # ggplot(data = world) +
# #     geom_sf(aes(fill = beer_servings)) +
# #     scale_fill_viridis_c(option = "plasma")
# # 
# # mapped_data <- joinCountryData2Map(beer_servings, joinCode = "ISO3", 
# #     nameJoinColumn = "country")
```



