---
title: 'Employee Data'
author: "Josephine Haag"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    highlight: zenburn
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---



```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(patchwork)
library(ggplot2)

```

# IBM HR Analytics


```{r loading_HR_dataset, include=FALSE}

hr_dataset <- read_csv( "datasets_1067_1925_WA_Fn-UseC_-HR-Employee-Attrition.csv")
glimpse(hr_dataset)

```

```{r cleaning_HR_dataset}

hr_cleaned <- hr_dataset %>% 
  clean_names() %>% 
  mutate(
    education = case_when(
      education == 1 ~ "Below College",
      education == 2 ~ "College",
      education == 3 ~ "Bachelor",
      education == 4 ~ "Master",
      education == 5 ~ "Doctor"
    ),
    environment_satisfaction = case_when(
      environment_satisfaction == 1 ~ "Low",
      environment_satisfaction == 2 ~ "Medium",
      environment_satisfaction == 3 ~ "High",
      environment_satisfaction == 4 ~ "Very High"
    ),
    job_satisfaction = case_when(
      job_satisfaction == 1 ~ "Low",
      job_satisfaction == 2 ~ "Medium",
      job_satisfaction == 3 ~ "High",
      job_satisfaction == 4 ~ "Very High"
    ),
    performance_rating = case_when(
      performance_rating == 1 ~ "Low",
      performance_rating == 2 ~ "Good",
      performance_rating == 3 ~ "Excellent",
      performance_rating == 4 ~ "Outstanding"
    ),
    work_life_balance = case_when(
      work_life_balance == 1 ~ "Bad",
      work_life_balance == 2 ~ "Good",
      work_life_balance == 3 ~ "Better",
      work_life_balance == 4 ~ "Best"
    )
  ) %>% 
  select(age, attrition, daily_rate, department,
         distance_from_home, education,
         gender, job_role,environment_satisfaction,
         job_satisfaction, marital_status,
         monthly_income, num_companies_worked, percent_salary_hike,
         performance_rating, total_working_years,
         work_life_balance, years_at_company,
         years_since_last_promotion)

```


1. How often do people leave the company (`attrition`)
```{r company_attrition_rate}
glimpse(hr_cleaned)


attrition_rate <- filter(hr_cleaned %>% 
         group_by(attrition) %>% 
         summarise(n = count(attrition)) %>% 
         mutate(attrition_rate = (n*100)/sum(n)), attrition == "Yes") #added column for the attrition rate %

attrition_rate

```

Answer:
The attrition rate is 16.1%.

2. How are `age`, `years_at_company`, `monthly_income` and `years_since_last_promotion` distributed? can you roughly guess which of these variables is closer to Normal just by looking at summary statistics? 

```{r HR_plots}
skim(hr_cleaned)

age_density <- ggplot(hr_cleaned, aes(age)) + 
  geom_density() + 
  labs(
    title = "Distribution of age",
    x = "Age"
  ) +
  theme_minimal() +
  NULL

years_at_company_density <- ggplot(hr_cleaned, aes(years_at_company)) + 
  geom_density() + 
  labs(
    title = "Distribution of years at company",
    x = "Years at company"
  ) +
  theme_minimal() +
  NULL

monthly_income_density <-ggplot(hr_cleaned, aes(monthly_income)) + 
  geom_density() +
  labs(
    title = "Distribution of monthly income",
    x = "Monthly income"
  ) +
  theme_minimal() +
  NULL

years_since_promotion <- ggplot(hr_cleaned, aes(years_since_last_promotion)) + 
  geom_density() +
  labs(
    title = "Distribution of years since promotion",
    x = "Years since promotion"
  ) +
  theme_minimal() +
  NULL

age_density
years_at_company_density
monthly_income_density
years_since_promotion


```


Answer:

We can look at how close the mean is to the 50th percentile and the relative positions of 25th and 75th percentiles are to the 50th percentile.`age` is close to normal distribution, while `years_at_company`, `monthly_income` and `years_since_last_promotion` are right screwed, with most small values in the left hand, a few large values in the right. 


3. How are `job_satisfaction` and `work_life_balance` distributed? Don't just report counts, but express categories as % of total

```{r job_sat_and_worklife_balance}
job_sat <- hr_cleaned %>% 
  group_by(job_satisfaction) %>% 
  summarise(n = count(job_satisfaction)) %>% 
  mutate(job_sat_perc = (n*100)/sum(n))

work_life_balance <- hr_cleaned %>% 
  group_by(work_life_balance) %>% 
  summarise(n = count(work_life_balance)) %>% 
  mutate(work_life_perc = (n*100)/sum(n)) %>% 
  arrange(desc(work_life_perc))
# 
# kbl(job_sat) %>% 
#   kable_styling(bootstrap_options=c("striped","hover","condensed","responsive"))
# kbl(work_life_balance) %>% 
#   kable_styling(bootstrap_options=c("striped","hover","condensed","responsive"))

```

Answer:

`job_satisfaction` is right skewed, with around 60% of the values on the right.
`work_life_balance` seems to better resemble a normal distribution, a bit skewed to the right.


4. Is there any relationship between monthly income and education? Monthly income and gender?
```{r monthly_income_gender_education_plots, fig.width=20}
library(hrbrthemes)
options("scipen"=100, "digits"=4)

monthly_income_education_plot <- ggplot(hr_cleaned,
   aes(
     x = education, 
     y = monthly_income 
   )) + 
  geom_col() +
  theme_minimal() +
  theme(text = element_text(size=40))+
  labs(
    title = "Monthly income vs. education",
    x = "Monthly income",
    y = "Education"
  ) +
  NULL

monthly_income_gender_plot <- ggplot(hr_cleaned,
   aes(
     x = gender, 
     y = monthly_income 
   )) + 
  geom_col() +
  theme_minimal() +
  theme(text = element_text(size=30))+
  labs(
    title = "Monthly income vs. gender",
    x = "Monthly income",
    y = "Gender"
  ) +
  NULL

monthly_income_education_plot
monthly_income_gender_plot

```

Answer:

- There is an inverted U-shaped relationship between education and monthly income. Bachelors' income is higher than college's and under college's, since most of the high-payment jobs require at least bachelor's degrees. And Bachelors' income is also greater than masters' and PhDs' income, probably because given they all meet most companies' basic education requirement, bachelors always have more working experience since they step into careers earlier.
- Male monthly incomes are higher than female. Gender inequality is still very common in job markets.


5. Plot a boxplot of income vs job role. Make sure the highest-paid job roles appear first
```{r income_jobrole_boxplot, fig.width=40, fig.height=20}

income_jobrole_boxplot <- ggplot(hr_cleaned, 
       aes(
         x= reorder(job_role, -monthly_income), 
         y = monthly_income)) + 
  geom_boxplot() +
  labs(
    title = "Income per job role",
    x = "Job role",
    y = "Income"
  ) +
  theme_minimal() +
  theme(text = element_text(size=30))+
  NULL

income_jobrole_boxplot

```


6. Calculate and plot a bar chart of the mean (or median?) income by education level.

```{r barchart_income_education}
Education_mean_income_bar <- ggplot(hr_cleaned, aes(
    x = education, 
    y = mean(monthly_income
    ))) + 
  geom_col() +
  labs(
    title = "Mean monthly income per education",
    x = "Education",
    y = "Mean monthly income"
  ) +
  theme_minimal() +
  theme(text = element_text(size=10))+
  NULL

Education_median_income_bar <- ggplot(hr_cleaned, aes(
    x = education, 
    y = median(monthly_income
    ))) + 
  geom_col() +
  labs(
    title = "Median monthly income per education",
    x = "Education",
    y = "Median monthly income") +
  theme_minimal() +
  theme(text = element_text(size=10))+
  NULL

Education_mean_income_bar
Education_median_income_bar

```

7. Plot the distribution of income by education level. Use a facet_wrap and a theme from `ggthemes`

```{r distribution_by_education, fig.width=10}

Monthly_income_education_distr <- ggplot(hr_cleaned, aes(
    x = monthly_income
  )) + 
  geom_histogram() + 
  facet_wrap(~education,ncol=2) + 
  labs(
    title = "Distribution of income by education",
    x = "Education"
  ) +
  theme_economist_white() +
  NULL

Monthly_income_education_distr

```


1. Plot income vs age, faceted by `job_role`
```{r Income_vs_age_by_jobrole}

Income_vs_age_facet <- ggplot(hr_cleaned, aes(
    x = age, 
    y = monthly_income
  )) + 
  geom_col(width = 2) + 
  facet_wrap(~job_role) +
  labs(
    title = "Income vs. age by job role",
    x = "Age",
    y = "Monthly income"
  ) +
  theme_minimal() +
  NULL

Income_vs_age_facet

```
