---
title: 'Social Media'
author: "Josephine Haag"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    highlight: default
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---
# SOCIAL MEDIA

```{r load-libraries, echo=FALSE, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(plotly)
library(ggthemr)
ggthemr('dust')
```

```{r, setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)
# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
library(ggthemr)
ggthemr('dust')

if (!requireNamespace("devtools", quietly = TRUE))
  install.packages("devtools")

devtools::install_github("calligross/ggthemeassist")
```


The [General Social Survey (GSS)](http://www.gss.norc.org/) gathers data on American society in order to monitor and explain trends in attitudes, behaviours, and attributes. Many trends have been tracked for decades, so one can see the evolution of attitudes, etc in American Society.

In the following, I will look into the users of different social medias in the US. 

```{r, read_gss_data, cache=TRUE, include=FALSE}
gss <- read_csv(here::here("content", "blogs", "smallgss2016.csv"), 
                na = c("", "Don't know",
                       "No answer", "Not applicable"))
glimpse(gss)
```

## Instagram and Snapchat, by sex

Can we estimate the *population* proportion of Snapchat or Instagram users in 2016?

```{r}
gss_snap_insta <- gss %>%
  # create the new variable for snap_insta
  mutate(snap_insta = case_when(
    snapchat == "Yes" | instagrm == "Yes" ~ "Yes", 
    snapchat == "No" & instagrm == "No" ~ "No",
    TRUE ~ NA_character_)) 
  
# calculate the proportion
gss_prop <- gss_snap_insta %>%
  filter(!is.na(snap_insta)) %>% 
  group_by(snap_insta) %>% 
  summarise(count = n()) %>% 
  mutate(prop = count/sum(count)) 
gss_prop

# percentile_ci
snap_insta_formula_ci <- gss_snap_insta %>%
  group_by(sex) %>%
  # using the CI formula m +- z * se
  summarize(p_hat = prop(snap_insta == "Yes"),
            count = n(),
            se = sqrt((p_hat * (1 - p_hat)) / count),
            t_critical = qt(0.975, count-1),
            margin_of_error = t_critical*se,
            lower_95 = p_hat - margin_of_error,
            upper_95 = p_hat + margin_of_error)
snap_insta_formula_ci

# additional plot for practice
histogram_si <- ggplot(snap_insta_formula_ci,  
       aes(x = sex,
           y = p_hat,
           colour = sex)) +
   geom_errorbar(aes(ymin = lower_95,
                     ymax = upper_95),
                     width = 0.05,
                     size  = 0.5) +
   geom_point(shape = 15,
              size  = 4) +
   scale_y_continuous(limits=c(0,1)) +
   theme(axis.title   = element_text(face  = "bold")) +
   labs(x = "Gender",
     y = "",
   title = "Women are more likely than men to use Snapchat or Instagram in 2016") +
   theme(legend.position = 'none') + 
   NULL
histogram_si
```


## Twitter, by education level

Can we estimate the *population* proportion of Twitter users by education level in 2016?. 

```{r}
# turn degree from a character variable into a factor variable 
gss_factor <- gss %>%
  mutate(degree = factor(degree, order = TRUE, levels = c("Lt high school", "High school", "Junior college", "Bachelor", "Graduate")))

# create the new variable bachelor_graduate and assign values based on criteria below:
gss_degree <- gss_factor %>%
  mutate(bachelor_graduate = case_when(
    degree %in% c("Bachelor","Master") ~ "Yes", 
    degree %in% c("Lt high school", "High school", "Junior college") ~ "No",
    TRUE ~ "NA"
    )) 
# filter out NAs and filter for bachelor and graduates
gss_bachelor <- gss_degree %>%
  filter(bachelor_graduate == "Yes")

# calculate the proportion of `bachelor_graduate` using twitter
gss_twitter_prop <- gss_bachelor %>%
  group_by(twitter) %>%
  filter(twitter != "NA") %>% 
  summarize(n = n()) %>%
  mutate(prop_t = n/sum(n))
  
gss_twitter_prop

# construct CI for `bachelor_graduate` (not) using twitter (but with the NAs in the 100 % )
twitter_formula_ci <- gss_twitter_prop %>%
  filter(twitter != "NA") %>%
   # use the CI formula m +- z * se
  summarize(twitter = twitter,
            p_hat_t = prop_t,
            # se for difference in proportions is sqrt(p(1-p)/n)
            se_t = sqrt((p_hat_t *(1-p_hat_t))/n),
            count_t = n,
            t_critical_t = qt(0.975, count_t-1),
            margin_of_error_t = t_critical_t*se_t,
            lower_95_t = p_hat_t - margin_of_error_t,
            upper_95_t = p_hat_t + margin_of_error_t)
twitter_formula_ci
  
# check if Confidence Intervals overlap
barchart_t <- ggplot(twitter_formula_ci,  
       aes(x = twitter,
           y = p_hat_t,
           colour = "#1DA1F2")) +
   geom_errorbar(aes(ymin = lower_95_t,
                     ymax = upper_95_t),
                     width = 0.05,
                     size  = 0.5) +
   geom_point(shape = 15,
              size  = 4) +
   scale_y_continuous(limits=c(0,1)) +
   theme(axis.title = element_text(face  = "bold")) +
   labs(x = "Twitter usage",
        y = "",
   title = "Twitter usage among people with a bachelor or graduate degree in 2016") +
   theme(legend.position = 'none') + 
   NULL

barchart_t
```


## Email usage

Can we estimate the population parameter on time spent on email weekly?

```{r}
# create new variable email
gss_email <- gss %>% 
  filter(emailmin != "NA", emailhr != "NA") %>%
  mutate(emailmin = as.numeric(emailmin), emailhr = as.numeric(emailhr)) %>%
  mutate(email = emailhr*60 + emailmin)

# visualise the distribution
plot_email <- ggplot(gss_email, 
  aes(x = email)) + 
  geom_density() + 
  geom_vline(aes(xintercept=mean(email, na.rm = T)),   # Ignore NA values for mean
               color="red", linetype="dashed", size=1) +
  geom_vline(aes(xintercept=median(email, na.rm = T)),   # Ignore NA values for mean
               color="blue", linetype="dashed", size=1) +
  annotate("text", x = 800, y = 0.0025, label = "mean", color="red") +
  annotate("text", x = 500, y = 0.00261, label = "median", color="blue") +
  theme_bw()+
  theme(legend.position = "none") +
  labs(
    title = "Right-skewed distribution of time spent on emails by US adults",
    x = "Minutes spent on emails per week",
    y = " ") +
  NULL

plot_email

# calculate a 95% bootstrap confidence interval for the mean
boot_email <- gss_email %>% 
  specify(response = email) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean") 
  
boot_email

# 95% CI and humanize units
percentile_ci <- boot_email %>%
  get_confidence_interval(level = 0.95, type = "percentile") %>%
  mutate(lower_ci_hum = lower_ci %/% 60 + (lower_ci %% 60)/100, upper_ci_hum = upper_ci %/% 60 + (upper_ci %% 60)/100)

percentile_ci
visualise(boot_email)
```